The pollination trade-off

### 2019-03-17 paper/manuscript.Rmd
Basic spellcheck

However, in natural communities, plants rarely share all available pollinators.
One explanation is that plants are effectively competing for mutualistic partners because there is a *trade-off* between the benefits gained from maximising the number of mutualistic partners and the costs of sharing them with other plant species [@waser_interspecific_1978]. 
It has been clear for a long time [both experimentally and theoretically; @robertson_philosophy_1895; @lewis_experimental_1961; @levin_competition_1970] that when competition for pollination is strong enough, it can be detrimental to plants' fitness. 
In fact, it can be sufficient to drive ecological differentiation [for example by driving the evolution of of both traits and reproductive strategies; @mitchell_new_2009; @caruso_competition_2000]. 
It is less clear, however, how exactly ecological differentiation, in turn, affect the trade-off between facilitation and competition that is involved in the pollination service.

There are two main mechanisms through which competition for pollinators can affect plant fertilization [@morales_interspecific_2008]. 
The first is intraspecific pollen transfer. 
This happens, for example, when plants with distinctively attractive flowers might reduce the number of visits---and hence the amount of *conspecific pollen* deposited by animals---to those less attractive neighbouring plants [@yang_pollinator_2011]. 
The second is intraspecific pollen transfer and occurs when plants share pollination partners. 
In that case, even receiving a visit might not necessarly translate into fertilization [@campbell_mechanism_1985] because a focal plant might receive *heterospecific pollen* or pollen from the focal plant might be lost to different species. 
Generally speaking, the higher the amount of conspecific pollen (both in absulute terms and relative to heterospecific pollen) the better the pollination service received by the focal plant.
These two mechanisms of competition, by definition, occur at the community scale. 
However, with few exceptions [@rathcke_interactions_1988; @lopezaraizamikel_impact_2007; @tur_evaluating_2016; @hegland_relative_2009; @aizen_reproductive_2010], most of what we know about the deposition of conspecific and heterospecific pollen---and its relationship to competition vs. facilitation in pollination systems---is based on studies of two plant species.

At the community scale the factors that determine the patterns of conspecific and heterospecific pollen deposition are tightly intertwined, operate simultaneously, and may lead to emergent phenomena not observed at smaller scales [@flanagan_effects_2011]. 
For instance, recent empirical evidence suggests that plants with flowering traits that are "original" relative to others in the community generally have fewer interaction partners [@coux_linking_2016]. 
This evidence is aligned with the notion that a species that interacts with few species does so strongly with each of them whereas a species that interacts with a large number of species does so comparatively weakly (TODO: ref needed). 
If evolutionary specialisation occurs by changing traits to focus on fewer but better partners [@caruso_competition_2000], we should expect a reduction of competition for pollinators in species with "original" traits and an increase in species with a large number of interaction partners [@gibson_can_2012]. 
Alternatively, it might also be the case that density (for example in terms of flower or pollen counts) is the dominant force driving pollen transfer [@seifan_outcome_2014]. 
Abundant plant species might experience a dilution of available pollinators [@feinsinger_effects_1987] but might also receive more effective visits by capitalising on a larger share of both visits and pollinator "real state". 
In this case, a potential reduction the absolute amount of conspecific pollen received could be compensated by an increase amount relative to heterospecific pollen. 
Altogether, it is clear that some ecological factors can influence pollen deposition and that these factors, in the specific context of pollination, could be used to roughly define a species *realised niche* in their community. 
However, the intertwined nature of these factors makes it hard to determine wether they favour or hinder the pollination service at the community level. 
Moreover, their relative importance has not yet been empirically evaluated in natural plant communities.

Here, we investigate competition for pollination at the community scale using empirical data from eleven plant-pollinator communities in the Argentinian Pampas. 
**First**, we estimate the extent to which the net outcome of plants sharing polinators is facilitation or competition. 
If facilitation is the primary outcome, we could expect two tings. 
On the one hand, we could expect an overall positive relationship between the amount of conspecific pollen deposited in stigmas and the amount of heterospecific pollen [@tur_evaluating_2016]. 
The higher the deposition of conspecific pollen relative to heterospecific, the larger the slope of that relationship.
On the other, we could also expect a larger amount of conspecific pollen deposited in stigmas when flowers are open to animal pollinators than is deposited due to self-pollination. 
**Second**, we investigate the relative contribution that four ecological factors have on the pollination service. 
Sepecifically, we hypothesize that competition (measured as absolute and relative conspecific pollen deposition) should increase for plants that share many pollination partners.
However, we also hypothesize that other factors like the plant's functional originality, its relative abundance, and the visit effectiveness should have the potential to compensate for this increase in competition. 
**Third**, we examine how community context influences the way plants use these factors to minimise competition. 
Although some plant species might be widespread across multiple locations, competition for pollinators occurs at the community level and hence we predict the competition for pollination niche to be influenced by the local context. 
If plant strategy is not flexible across communities (TODO: ref needed), we could expect niches to be more similar within species than across species in the study. 
On the contrary, if a species' strategy is flexible, we should expect plants to be able to occupy different niches in each community such that competition could be minimised regardless of the context.

@@

We collected data from eleven co-flowering plant communities and their pollinators in three locations in the Argentinian Pampas. 
These locations are distributed across a precipitation gradient with a humid temperate climate in the eastern Flooding Pampas and becoming drier as one moves inland and westward towards the Pampean Grasslands. 
This precipitation gradient generatesd a natural gradient of network architecture which allowed us to examine the generality of our findings. 
In each location, we sampled two restored and two agricultural fragments, except in the Flooding Pampas, where we were only able to sample one restored fragment due to the lack of available sites.

@@

Our first objective was to identify whether the outcome of plants sharing polinators at the community level is facilitation or competition. 
The first line of evidence can be found via the relationship between heterospecific and conspecific pollen deposited in plant stigmas [@tur_evaluating_2016]. 
This is because heterospecific pollen can be seen as an unavoidable cost that plants must pay for receiving conspecific pollen from shared pollinators. 
A negative relationship (when shared pollinators increase heterospecific pollen loads and decrease conspecific loads) may indicate that competition for pollination is strong. 
This negative relationship could arise, for example, when another, more dominant species, monopolizes a limited number of pollinators [@mitchell_new_2009; @lopezaraizamikel_impact_2007], when heterospecific pollen deposited in the stigma interferes with deposition of conspecific pollen [@runquist_pollinator-mediated_2012], or when conspecific pollen is lost to heterospecific stigmas [@mitchell_new_2009; @muchhala_interspecific_2012]. 
In contrast, a positive relationship is what would be observed when facilitation is the dominant outcome. 
For example, when the abundance of shared pollinators is positively influenced by floral resources, these positive effects ripple through the mutualistic network in a chain of positive feedback loops [@liao_presence_2011].

@@

To evaluate the relationship between heterospecific and conspecific pollen, we used a generalised linear mixed model (GLMM) with a Poisson error distribution [@tur_evaluating_2016]. 
In this model, the number conspecific pollen grains per stigma was the response variable and the number of heterospecific grains the predictor (the number of stigmas sampled from the flower was added as an offset term). 
The slope and intercept of this relationship were allowed to differ between plant species within each community, which were modelled as a random effect. 
Models were fitted using the function `glmer` from the R package `lme4 1.1-19` [@bates_fitting_2015]. 
To compare the conspecific pollen deposited with and without animal-mediated pollination, we followed the same approach as above. 
In this case, however, the predictor was the treatment (wether the flower was bagged or unbagged) instead of the number of heterospecific grains.

@@

We investigated the impact of the ecological factors on pollination services.
Generally speaking, the more conspecific pollen deposited in stigmas, both absolutely and relative to heterospecific pollen, the better the pollination service. 
Specifically, we used two sets of linear mixed models (LMM) in which the response variables are the number of conspecific and heterospecific pollen grains deposited per stigma in flowers open to animal-mediated pollination. 
In general, using LMMs in which pollen loads were log-transformed (Figure \@ref(S-fig:fig-deposition-gain-log-transformed-or-not)) offered better fit than a GLMM with Poisson (or quasipoisson) error structure. 
Models were fitted using the R package `nlme 3.1-131` [@pinheiro_nlme_2018].

@@

To estimate the coefficients and to perform model selection and quantify the associated uncertainty, we used a combination of multi-model inference and bootstrap resampling with one hundred replicates. 
First, we performed model selection using the AICc and determined the likelihood of each candidate model (a particular combination of predicting variables) by calculating the median \(\Delta\)AICc (relative to the most likely model) for each bootstrap sample. 
As we wanted model coefficients from more likely candidate models to carry more weight in our results, we sampled the coefficients for our variables proportional to the likelihood of its candidate model. 
We "shrank" the estimated coefficient of potentially unimportant variables by setting their estimate to 0 when the variable was not included in the model. 
Finally, we used these distributions of the model coefficients to estimate their mean impact on the pollination service (in terms of absoulte and relative conspecific pollen deposition). 
Again, here we used bootstrap resampling to estimate the uncertainty around these means.

@@

Our last objective was to tease apart the plant strategies that might reduce competition and, importantly, how the community context influences these strategies. 
If community context plays a relatively small role or the species strategy is inflexible we would expect plants of the same species to fill similar competition for pollination niches across different communities. 
Alternatively, if the community plays a large role and plants sttrategies are flexible, we should be able to observe differences in the niche a plant species occupies across communities. 
First, we used a principal component analysis (PCA) of the four ecological variables. 
We scaled variables across the whole study to ensure that the PCA space does not change acording to the species present in each community. 
We define a species' niche in a community as its coordinates in PCA space. 
For each species that was present in two or more communities, we then calculated (*i*) the median distance between the strategies that the species uses in different communities and (*ii*) the area of the convex hull defined by these points in the first two principal components (only for species present in three or more communities). 
We then compared the magnitude of these two metrics to those obtained with 99 Monte Carlo randomizations in which we replaced the strategy of the focal plant species by that of another randomly selected species in the dataset.

@@

We first examined the degree of facilitation/competition in natural communities. Specifically, we examined the slope of the relationship between the heterospecific and conspecific pollen deposited on stigmas of each species in a community. 
We found that overall neither facilitation nor competition dominates (Table \@ref(S-tab:table-facilitation-models-general)). 
Indeed, we found that the proportion of species that experienced a significantly positive or neggative relationship between heterospecific and conspecific pollen than neggative was very similar (`r counts_facilitation_slope$np[3]`% and `r counts_facilitation_slope$np[1]`%, respectively; Figure \@ref(fig:fig-trade-off-predictions)a; Table \@ref(S-tab:table-facilitation-models-random)). 
Most of the variation between plants in a facilitation-competition gradient was driven by differences *between* species; however, we also observed some important differences *within* species. 
For instance, `r numbers2words(nrow(sp_both_pos_neg))` of the species sampled across multiple communities (`r sp_both_pos_neg_names`) showed evidence that, the same species, can experience facilitation (positive relationships) in some communities and competition (negative relationships) in others. 
Interestingly, although there was a wide variation in the slope of the relationship between heterospecific and conspecific pollen almost all plants had more conspecific than heterospecific pollen deposited on their stigmas (Figure \@ref(fig:fig-trade-off-predictions)b).

@@

Of the four variables we considered, we found that a plant's mean visit effectiveness and abundance were the most important ecological variables predicting pollen deposition in plant stigmas (Figure \@ref(fig:fig-average-qual-quant)a). 
Surprisingly, the number of shared pollinators was comparatively unimportant, particularly for models of heterospecific pollen deposition, as it was only ever included in models with relatively large AICc values (Table \@ref(S-tab:table-aic-fixed-models)).

We found that the relationship between each of the ecological factors and pollen deposition was similar for both conspecific and heterospecific pollen. 
That is, strategies that were associated with an increase in conspecific pollen were also associated with an increase in heterospecific pollen deposition. 
Specifically, the plants' mean visit effectiveness had a positive effect on pollen deposition (Figure \@ref(fig:fig-average-qual-quant)b). 
However, the effect size was slightly larger for heterospecific than for conspecific pollen. 
This indicates that, although there is a positive association between visit effectiveness and the *absolute* amount of conspecific pollen, there is a negative relationship with the *relative* amount of conspecific relative to heterospecific pollen (Figure \@ref(fig:fig-average-qual-quant)c). 
In contrast, a plants' relative abundance had a negative effect on the absolute amount of conspecific pollen, but the mean difference between the coefficients in the models indicates a positive association with the relatvie amount of conspecific pollen (Figure \@ref(fig:fig-average-qual-quant)c). 
The third most important variable, functional originality, had a positive, although comparatively smaller, association with both the absolute and relative amount of conspecific pollen deposited. 
Finally, the number of shared pollinators had a negative and neutral association with conspecific and heterospecific pollen, but these impacts were small when compared to the other variables.

Overall, collinearity did not influence our findings qualitatively.
Although relative abundance, the number of shared pollinators, and the visit effectiveness were all positively correlated (Figure \@ref(S-fig:fig-correlation)), the effect each had on conspecific pollen was similar among models that included all or just some of these three explanatory variables (Figure \@ref(S-fig:fig-model-results)). 
One exception was visit effectiveness, which exhibits a positive association with the relative amount of conspecific pollen under some variable combinations (all with relatively low AICc support).

@@

We used a PCA of the analysed species to investigate whether plant's niche of competition for pollination are similar across communities or whether they are flexible, and therefore a reflection of the community context. 
The first two PCA components explained `r round(this_pca$eig[2, 3])`% of the total variance (Figure \@ref(fig:fig-pca)a). 
The first component was dominated by changes in visit effectiveness and relative abundance while the second component was dominated by the number of shared pollinators and the plant's functional originality. 
When we locate the species that were sampled in more than one community in the first two PCA components (Figure \@ref(fig:fig-pca)b), we observe that the niches of any given species do not tend to be close to each other. 
Indeed, when we measured the median distance between the plants coordinates, we found that it was only significantly smaller than that of randomisations for only `r numbers2words(sum(median_distances < 0.05))` of the `r numbers2words(length(median_distances))` analysed species (Figure \@ref(fig:fig-pca)c).

@@

These trade-offs were most evident in precisely the factors that most strongly modulate pollen deposition. 
Specifically, we found that the factors related to the relative dominance of the species (in terms of pollen and visits, or abundance) also tended to be the most important predictors of pollen deposition. 
At the same time, receiving high visitation increases the absolute levels of conspecific pollen deposition but increasses heterospeicific deposition to a greater extent---even when the visitors are likely carry a high proportion of conspecific pollen. 
Similarly, being abundant reduces the amount of heterospecific pollen deposited but at the expense of a noticeable reduction in the amount of conspecific pollen. 
Multiple studies recognise the importance of visitation and abundance to competition for pollination [@feldman_when_2004; @munoz_presence_2008; @morales_interspecific_2008]. 
Our results strongly corroborate the importance of these two factors but also suggest that [because visitation, pollen production and abundance are correlated; @sargent_role_2006], plants must reach a fine balance between increasing conspecific pollen and reduce heterospecific pollen deposition to minimise competition for pollination.

Further evidence of the intricacies of balancing costs and benefits can be found when we examine the other two factors analysed in our models. 
We show that being a specialist and sharing no pollinators might be preferable in the short term (as it is generally detrimental to pollen deposition). 
However, over long periods of time there could be a risk associated to a specialist plant having no pollinators. 
Hence, is likely that, to ensure long term survival, plants also need to balance this risk with the costs of sharing pollinators. 
One possible solution is to share pollinators *and* have original traits---as we show that trait originallity is generally benefitial to pollen deposition and it’s commonly thought that species that are further from others in trait space suffer reduced competition. 
Yet again, there are two possible caveats to this strategy that highlight the interrelatedness of the ecological factors. 
First, in a mutualism context, it is also possible that trait originality could come at the cost of being less ‘apparent’ to pollinators. 
Second, the usuallly negative, relationship between originality and generalism has been shown depend on the plants abundance [@coux_linking_2016], with generalist species being able to also have original traits only when they are abundant enough to provide a valuable reward to make it worth to pollinators.

How precisely plants balance the costs and benefits in the short and the long term likely depends on the species sensitivity to different risks but also, as we show here, on the community it belongs to. 
For instance, a plant species that experience a large reduction in seed set due to heterospecific pollen deposition might be better off by sharing few pollinators, while sharing pollinators makes more sense for those mainly affected by low levels of conspecific pollen deposition [@campbell_mechanism_1985].
Nevertheless, the diversity of the facilitation/competition outcomes across communities we observed and the relevance of random effects in our models suggests that the community context has a very important effect on the way plant species respond to competition for pollination beyond species identity. 
Perhaps overgeneralising, some plant families (Asteraceae for example) are often generalists in their communities, others (Orchidaceae) are known to be more commonly specialised. 
While it is possible, from an evolutionary perspective, that selection for a particular strategy might say something about the community in which a species has typically inhabited during its evolutionary history, we show that most plant species are still flexible enough to adopt markedly different niches in different communities. 
From a more applied perspective, flowering plants are sometimes introduced to attract pollinators on other nearby plants. 
Our results highlight the difficulties involved with predicting wether the introduced plant species will facilitate or compete with neighbours. 
This is so because the outcome might depend not only on the particular niche each species occupies in a particular community, but also on other factors that we were unable to measure [e.g. pollinator behaviour and densities or spatial context; @ye_competition_2014; @flanagan_effects_2011; @cariveau_spatially_2009].
However, our results also suggest that introduced species are more likely to be benefitial if they have differing traits to local plants. 
<!-- cite melissa's paper if available before publication -->

Overall, our results suggest that pollination communities might not necessarily be the paradigmatic mutualistic communities we often believe. 
Instead, we confirm that plant competition is pervasive in pollination communities and that potential strategies to minimise competition are likely involve strong trade-offs in the pollination service either in the short or the long term. 
As expected, not sharing pollinators and being functionally unique reduces competition for pollination by increassing conspecific pollen deposition (both in absolute numbers and relative to heterospecific pollen). 
However, this increasse is surprisingly small when compared to the effect of relative abundance and visitation. 
Interestingly, these two factors also involve the largest trade-offs between increasing deposition of conspecific pollen and reducing that of heterospecific pollen. 
Despite large intra-species differences among species our results provide solid evidence of the role that competition for pollination plays in natural communities. 
Species coexistence is not solely determined by direct and indirect mutualistic relationships, but also strongly shaped by the multiple intewined ecological factors that shape competition for pollination at the community level. 
Achieving a better understanding how biodiversity is supported by pollination communities will require seeing them as both mutualistic and competitive comminities.

>>>>

However, in natural communities, plants rarely share all available pollinators.
One explanation is that plants are effectively competing for mutualistic partners because there is a *trade-off* between the benefits gained from maximising the number of mutualistic partners and the costs of sharing them with other plant species [@waser_interspecific_1978]. 
It has been clear for a long time [both experimentally and theoretically; @robertson_philosophy_1895; @lewis_experimental_1961; @levin_competition_1970] that when competition for pollination is strong enough, it can be detrimental to plants' fitness. 
In fact, it can be sufficient to drive ecological differentiation [for example by driving the evolution of both traits and reproductive strategies; @mitchell_new_2009; @caruso_competition_2000]. 
It is less clear, however, how exactly ecological differentiation, in turn, affect the trade-off between facilitation and competition that is involved in the pollination service.

There are two main mechanisms through which competition for pollinators can affect plant fertilization [@morales_interspecific_2008]. 
The first is intraspecific pollen transfer. 
This happens, for example, when plants with distinctively attractive flowers might reduce the number of visits---and hence the amount of *conspecific pollen* deposited by animals---to those less attractive neighbouring plants [@yang_pollinator_2011]. 
The second is intraspecific pollen transfer and occurs when plants share pollination partners. 
In that case, even receiving a visit might not necessarily translate into fertilization [@campbell_mechanism_1985] because a focal plant might receive *heterospecific pollen* or pollen from the focal plant might be lost to different species. 
Generally speaking, the higher the amount of conspecific pollen (both in absolute terms and relative to heterospecific pollen) the better the pollination service received by the focal plant.
These two mechanisms of competition, by definition, occur at the community scale. 
However, with few exceptions [@rathcke_interactions_1988; @lopezaraizamikel_impact_2007; @tur_evaluating_2016; @hegland_relative_2009; @aizen_reproductive_2010], most of what we know about the deposition of conspecific and heterospecific pollen---and its relationship to competition vs. facilitation in pollination systems---is based on studies of two plant species.

At the community scale, the factors that determine the patterns of conspecific and heterospecific pollen deposition are tightly intertwined, operate simultaneously, and may lead to emergent phenomena not observed at smaller scales [@flanagan_effects_2011]. 
For instance, recent empirical evidence suggests that plants with flowering traits that are "original" relative to others in the community generally have fewer interaction partners [@coux_linking_2016]. 
This evidence is aligned with the notion that a species that interacts with few species does so strongly with each of them whereas a species that interacts with a large number of species does so comparatively weakly (TODO: ref needed). 
If evolutionary specialisation occurs by changing traits to focus on fewer but better partners [@caruso_competition_2000], we should expect a reduction of competition for pollinators in species with "original" traits and an increase in species with a large number of interaction partners [@gibson_can_2012]. 
Alternatively, it might also be the case that density (for example in terms of flower or pollen counts) is the dominant force driving pollen transfer [@seifan_outcome_2014]. 
Abundant plant species might experience a dilution of available pollinators [@feinsinger_effects_1987] but might also receive more effective visits by capitalising on a larger share of both visits and pollinator "real state". 
In this case, a potential reduction in the absolute amount of conspecific pollen received could be compensated by an increase in the amount relative to heterospecific pollen. 
Altogether, it is clear that some ecological factors can influence pollen deposition and that these factors, in the specific context of pollination, could be used to roughly define a species *realised niche* in their community. 
However, the intertwined nature of these factors makes it hard to determine whether they favour or hinder the pollination service at the community level. 
Moreover, their relative importance has not yet been empirically evaluated in natural plant communities.

Here, we investigate competition for pollination at the community scale using empirical data from eleven plant-pollinator communities in the Argentinian Pampas. 
**First**, we estimate the extent to which the net outcome of plants sharing pollinators is facilitation or competition. 
If facilitation is the primary outcome, we could expect two things. 
On the one hand, we could expect an overall positive relationship between the amount of conspecific pollen deposited in stigmas and the amount of heterospecific pollen [@tur_evaluating_2016]. 
The higher the deposition of conspecific pollen relative to heterospecific, the larger the slope of that relationship.
On the other, we could also expect a larger amount of conspecific pollen deposited in stigmas when flowers are open to animal pollinators than is deposited due to self-pollination. 
**Second**, we investigate the relative contribution that four ecological factors have on the pollination service. 
Specifically, we hypothesize that competition (measured as absolute and relative conspecific pollen deposition) should increase for plants that share many pollination partners.
However, we also hypothesize that other factors like the plant's functional originality, its relative abundance, and the visit effectiveness should have the potential to compensate for this increase in competition. 
**Third**, we examine how community context influences the way plants use these factors to minimise competition. 
Although some plant species might be widespread across multiple locations, competition for pollinators occurs at the community level and hence we predict the competition for pollination niche to be influenced by the local context. 
If plant strategy is not flexible across communities (TODO: ref needed), we could expect niches to be more similar within species than across species in the study. 
On the contrary, if a species' strategy is flexible, we should expect plants to be able to occupy different niches in each community such that competition could be minimised regardless of the context.

@@

We collected data from eleven co-flowering plant communities and their pollinators in three locations in the Argentinian Pampas. 
These locations are distributed across a precipitation gradient with a humid temperate climate in the eastern Flooding Pampas and becoming drier as one moves inland and westward towards the Pampean Grasslands. 
This precipitation gradient generated a natural gradient of network architecture which allowed us to examine the generality of our findings. 
In each location, we sampled two restored and two agricultural fragments, except in the Flooding Pampas, where we were only able to sample one restored fragment due to the lack of available sites.

@@

Our first objective was to identify whether the outcome of plants sharing pollinators at the community level is facilitation or competition. 
The first line of evidence can be found via the relationship between heterospecific and conspecific pollen deposited in plant stigmas [@tur_evaluating_2016]. 
This is because heterospecific pollen can be seen as an unavoidable cost that plants must pay for receiving conspecific pollen from shared pollinators. 
A negative relationship (when shared pollinators increase heterospecific pollen loads and decrease conspecific loads) may indicate that competition for pollination is strong. 
This negative relationship could arise, for example, when another, more dominant species, monopolizes a limited number of pollinators [@mitchell_new_2009; @lopezaraizamikel_impact_2007], when heterospecific pollen deposited in the stigma interferes with deposition of conspecific pollen [@runquist_pollinator-mediated_2012], or when conspecific pollen is lost to heterospecific stigmas [@mitchell_new_2009; @muchhala_interspecific_2012]. 
In contrast, a positive relationship is what would be observed when facilitation is the dominant outcome. 
For example, when the abundance of shared pollinators is positively influenced by floral resources, these positive effects ripple through the mutualistic network in a chain of positive feedback loops [@liao_presence_2011].

@@

To evaluate the relationship between heterospecific and conspecific pollen, we used a generalised linear mixed model (GLMM) with a Poisson error distribution [@tur_evaluating_2016]. 
In this model, the number conspecific pollen grains per stigma was the response variable and the number of heterospecific grains the predictor (the number of stigmas sampled from the flower was added as an offset term). 
The slope and intercept of this relationship were allowed to differ between plant species within each community, which were modelled as a random effect. 
Models were fitted using the function `glmer` from the R package `lme4 1.1-19` [@bates_fitting_2015]. 
To compare the conspecific pollen deposited with and without animal-mediated pollination, we followed the same approach as above. 
In this case, however, the predictor was the treatment (whether the flower was bagged or unbagged) instead of the number of heterospecific grains.

@@

We investigated the impact of the ecological factors on pollination services.
Generally speaking, the more conspecific pollen deposited in stigmas, both absolutely and relative to heterospecific pollen, the better the pollination service. 
Specifically, we used two sets of linear mixed models (LMM) in which the response variables are the number of conspecific and heterospecific pollen grains deposited per stigma in flowers open to animal-mediated pollination. 
In general, using LMMs in which pollen loads were log-transformed (Figure \@ref(S-fig:fig-deposition-gain-log-transformed-or-not)) offered a better fit than a GLMM with Poisson (or quasipoisson) error structure. 
Models were fitted using the R package `nlme 3.1-131` [@pinheiro_nlme_2018].

@@

To estimate the coefficients and to perform model selection and quantify the associated uncertainty, we used a combination of multi-model inference and bootstrap resampling with one hundred replicates. 
First, we performed model selection using the AICc and determined the likelihood of each candidate model (a particular combination of predicting variables) by calculating the median \(\Delta\)AICc (relative to the most likely model) for each bootstrap sample. 
As we wanted model coefficients from more likely candidate models to carry more weight in our results, we sampled the coefficients for our variables proportional to the likelihood of its candidate model. 
We "shrank" the estimated coefficient of potentially unimportant variables by setting their estimate to 0 when the variable was not included in the model. 
Finally, we used these distributions of the model coefficients to estimate their mean impact on the pollination service (in terms of absolute and relative conspecific pollen deposition). 
Again, here we used bootstrap resampling to estimate the uncertainty around these means.

@@

Our last objective was to tease apart the plant strategies that might reduce competition and, importantly, how the community context influences these strategies. 
If community context plays a relatively small role or the species strategy is inflexible we would expect plants of the same species to fill similar competition for pollination niches across different communities. 
Alternatively, if the community plays a large role and plants strategies are flexible, we should be able to observe differences in the niche a plant species occupies across communities. 
First, we used a principal component analysis (PCA) of the four ecological variables. 
We scaled variables across the whole study to ensure that the PCA space does not change according to the species present in each community. 
We define a species' niche in a community as its coordinates in PCA space. 
For each species that was present in two or more communities, we then calculated (*i*) the median distance between the strategies that the species uses in different communities and (*ii*) the area of the convex hull defined by these points in the first two principal components (only for species present in three or more communities). 
We then compared the magnitude of these two metrics to those obtained with 99 Monte Carlo randomizations in which we replaced the strategy of the focal plant species by that of another randomly selected species in the dataset.

@@

We first examined the degree of facilitation/competition in natural communities. Specifically, we examined the slope of the relationship between the heterospecific and conspecific pollen deposited on stigmas of each species in a community. 
We found that overall neither facilitation nor competition dominates (Table \@ref(S-tab:table-facilitation-models-general)). 
Indeed, we found that the proportion of species that experienced a significantly positive or neggative relationship between heterospecific and conspecific pollen than neggative was very similar (`r counts_facilitation_slope$np[3]`% and `r counts_facilitation_slope$np[1]`%, respectively; Figure \@ref(fig:fig-trade-off-predictions)a; Table \@ref(S-tab:table-facilitation-models-random)). 
Most of the variation between plants in a facilitation-competition gradient was driven by differences *between* species; however, we also observed some important differences *within* species. 
For instance, `r numbers2words(nrow(sp_both_pos_neg))` of the species sampled across multiple communities (`r sp_both_pos_neg_names`) showed evidence that the same species, can experience facilitation (positive relationships) in some communities and competition (negative relationships) in others. 
Interestingly, although there was a wide variation in the slope of the relationship between heterospecific and conspecific pollen almost all plants had more conspecific than heterospecific pollen deposited on their stigmas (Figure \@ref(fig:fig-trade-off-predictions)b).

@@

Of the four variables, we considered, we found that a plant's mean visit effectiveness and abundance were the most important ecological variables predicting pollen deposition in plant stigmas (Figure \@ref(fig:fig-average-qual-quant)a). 
Surprisingly, the number of shared pollinators was comparatively unimportant, particularly for models of heterospecific pollen deposition, as it was only ever included in models with relatively large AICc values (Table \@ref(S-tab:table-aic-fixed-models)).

We found that the relationship between each of the ecological factors and pollen deposition was similar for both conspecific and heterospecific pollen. 
That is, strategies that were associated with an increase in conspecific pollen were also associated with an increase in heterospecific pollen deposition. 
Specifically, the plants' mean visit effectiveness had a positive effect on pollen deposition (Figure \@ref(fig:fig-average-qual-quant)b). 
However, the effect size was slightly larger for heterospecific than for conspecific pollen. 
This indicates that, although there is a positive association between visit effectiveness and the *absolute* amount of conspecific pollen, there is a negative relationship with the *relative* amount of conspecific relative to heterospecific pollen (Figure \@ref(fig:fig-average-qual-quant)c). 
In contrast, a plants' relative abundance had a negative effect on the absolute amount of conspecific pollen, but the mean difference between the coefficients in the models indicates a positive association with the relative amount of conspecific pollen (Figure \@ref(fig:fig-average-qual-quant)c). 
The third most important variable, functional originality, had a positive, although comparatively smaller, association with both the absolute and relative amount of conspecific pollen deposited. 
Finally, the number of shared pollinators had a negative and neutral association with conspecific and heterospecific pollen, but these impacts were small when compared to the other variables.

Overall, collinearity did not influence our findings qualitatively.
Although relative abundance, the number of shared pollinators, and the visit effectiveness were all positively correlated (Figure \@ref(S-fig:fig-correlation)), the effect each had on conspecific pollen was similar among models that included all or just some of these three explanatory variables (Figure \@ref(S-fig:fig-model-results)). 
One exception was visit-effectiveness, which exhibits a positive association with the relative amount of conspecific pollen under some variable combinations (all with relatively low AICc support).

@@

We used a PCA of the analysed species to investigate whether plants' niche of competition for pollination is similar across communities or whether they are flexible, and therefore a reflection of the community context. 
The first two PCA components explained `r round(this_pca$eig[2, 3])`% of the total variance (Figure \@ref(fig:fig-pca)a). 
The first component was dominated by changes in visit effectiveness and relative abundance while the second component was dominated by the number of shared pollinators and the plant's functional originality. 
When we locate the species that were sampled in more than one community in the first two PCA components (Figure \@ref(fig:fig-pca)b), we observe that the niches of any given species do not tend to be close to each other. 
Indeed, when we measured the median distance between the plants coordinates, we found that it was only significantly smaller than that of randomisations for only `r numbers2words(sum(median_distances < 0.05))` of the `r numbers2words(length(median_distances))` analysed species (Figure \@ref(fig:fig-pca)c).

@@

These trade-offs were most evident in precisely the factors that most strongly modulate pollen deposition. 
Specifically, we found that the factors related to the relative dominance of the species (in terms of pollen and visits, or abundance) also tended to be the most important predictors of pollen deposition. 
At the same time, receiving high visitation increases the absolute levels of conspecific pollen deposition but increases heterospecific deposition to a greater extent---even when the visitors are likely to carry a high proportion of conspecific pollen. 
Similarly, being abundant reduces the amount of heterospecific pollen deposited but at the expense of a noticeable reduction in the amount of conspecific pollen. 
Multiple studies recognise the importance of visitation and abundance to competition for pollination [@feldman_when_2004; @munoz_presence_2008; @morales_interspecific_2008]. 
Our results strongly corroborate the importance of these two factors but also suggest that [because visitation, pollen production and abundance are correlated; @sargent_role_2006], plants must reach a fine balance between increasing conspecific pollen and reduce heterospecific pollen deposition to minimise competition for pollination.

Further evidence of the intricacies of balancing costs and benefits can be found when we examine the other two factors analysed in our models. 
We show that being a specialist and sharing no pollinators might be preferable in the short term (as it is generally detrimental to pollen deposition). 
However, over long periods of time, there could be a risk associated with a specialist plant having no pollinators. 
Hence, is likely that to ensure long term survival, plants also need to balance this risk with the costs of sharing pollinators. 
One possible solution is to share pollinators *and* have original traits---as we show that trait originality is generally beneficial to pollen deposition and it’s commonly thought that species that are further from others in trait space suffer reduced competition. 
Yet again, there are two possible caveats to this strategy that highlight the interrelatedness of the ecological factors. 
First, in a mutualism context, it is also possible that trait originality could come at the cost of being less ‘apparent’ to pollinators. 
Second, the usually negative, relationship between originality and generalism has been shown depend on the plant's abundance [@coux_linking_2016], with generalist species being able to also have original traits only when they are abundant enough to provide a valuable reward to make it worth to pollinators.

How precisely plants balance the costs and benefits in the short and the long term likely depends on the species sensitivity to different risks but also, as we show here, on the community it belongs to. 
For instance, a plant species that experience a large reduction in seed set due to heterospecific pollen deposition might be better off by sharing few pollinators, while sharing pollinators makes more sense for those mainly affected by low levels of conspecific pollen deposition [@campbell_mechanism_1985].
Nevertheless, the diversity of the facilitation/competition outcomes across communities we observed and the relevance of random effects in our models suggests that the community context has a very important effect on the way plant species respond to competition for pollination beyond species identity. 
Perhaps overgeneralising, some plant families (Asteraceae for example) are often generalists in their communities, others (Orchidaceae) are known to be more commonly specialised. 
While it is possible, from an evolutionary perspective, that selection for a particular strategy might say something about the community in which a species has typically inhabited during its evolutionary history, we show that most plant species are still flexible enough to adopt markedly different niches in different communities. 
From a more applied perspective, flowering plants are sometimes introduced to attract pollinators on other nearby plants. 
Our results highlight the difficulties involved in predicting whether the introduced plant species will facilitate or compete with neighbours. 
This is so because the outcome might depend not only on the particular niche each species occupies in a particular community but also on other factors that we were unable to measure [e.g. pollinator behaviour and densities or spatial context; @ye_competition_2014; @flanagan_effects_2011; @cariveau_spatially_2009].
However, our results also suggest that introduced species are more likely to be beneficial if they have different traits to local plants. 
<!-- cite melissa's paper if available before publication -->

Overall, our results suggest that pollination communities might not necessarily be the paradigmatic mutualistic communities we often believe. 
Instead, we confirm that plant competition is pervasive in pollination communities and that potential strategies to minimise competition are likely to involve strong trade-offs in the pollination service either in the short or the long term. 
As expected, not sharing pollinators and being functionally unique reduces competition for pollination by increasing conspecific pollen deposition (both in absolute numbers and relative to heterospecific pollen). 
However, this increase is surprisingly small when compared to the effect of relative abundance and visitation. 
Interestingly, these two factors also involve the largest trade-offs between increasing deposition of conspecific pollen and reducing that of heterospecific pollen. 
Despite large intra-species differences among species our results provide solid evidence of the role that competition for pollination plays in natural communities. 
Species coexistence is not solely determined by direct and indirect mutualistic relationships, but also strongly shaped by the multiple intertwined ecological factors that shape competition for pollination at the community level. 
Achieving a better understanding of how biodiversity is supported by pollination communities will require seeing them as both mutualistic and competitive communities.
